---
title: "city_redwood"
author: "Kate Ham"
date: "2/14/2020"
output: 
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Libraries
library(tidyverse)
library(sf)
library(readxl) # part of the tidyverse


# Files
acs_rwc_file <- "C:/Users/katea/GitHub/dcl-2020-01/kate/c01-own/data/acs_rwc.rds"
boundaries_cbg_city_redwood_file <- "C:/Users/katea/GitHub/dcl-2020-01/kate/c01-own/data/boundaries_cbg_city_redwood.rds"
boundaries_city_redwood_file <- "C:/Users/katea/GitHub/dcl-2020-01/kate/c01-own/data/boundaries_city_redwood.rds"
bayarea_water_shp <- "C:/Users/katea/GitHub/dcl-2020-01/kate/c01-own/data-raw/stanford-qh320kj0191-shapefile/qh320kj0191.shp"
cl_city_redwood_file <- "C:/Users/katea/GitHub/dcl-2020-01/kate/c01-own/data/cl_city_redwood.rds"
xlsx_file <- "C:/Users/katea/GitHub/dcl-2020-01/kate/c01-own/data-raw/Bulk_Tax_Current_Layout w parcel level lat long_03102016.xlsx"

```

Read in relevant files.

```{r}
acs_rwc <- read_rds(acs_rwc_file)
boundaries_cbg_city_redwood <- read_rds(boundaries_cbg_city_redwood_file)
boundaries_city_redwood <- read_rds(boundaries_city_redwood_file) # EPSG 4269 datum NAD83
bayarea_water <- st_read(bayarea_water_shp) %>% st_transform(st_crs(boundaries_city_redwood))
cl_city_redwood <- read_rds(cl_city_redwood_file)

data_dic_luse <- 
  read_excel(xlsx_file, sheet = "LUSE") %>% 
  transmute(
    luse_code = as.double(`CdVal`),
    luse_desc = `CdDesc`
  )

# q1 <- 
#   file_data_dic %>% 
#   excel_sheets() %>% 
#   set_names() %>% 
#   map(read_xlsx, path = file_data_dic) %>% 
#   map(~ transmute(., code  = .$CdVal))

```

The ACS data includes estimates for	household income in the past 12 months (in 2018 inflation-adjusted dollars), from table B19001. This is the list of the variables:

* 'B19001_001' - Total
* 'B19001_002' - Less than $10 000
* 'B19001_003' - $10 000 to $14 999
* 'B19001_004' - $15 000 to $19 999
* 'B19001_005' - $20 000 to $24 999
* 'B19001_006' - $25 000 to $29 999
* 'B19001_007' - $30 000 to $34 999
* 'B19001_008' - $35 000 to $39 999
* 'B19001_009' - $40 000 to $44 999
* 'B19001_010' - $45 000 to $49 999
* 'B19001_011' - $50 000 to $59 999
* 'B19001_012' - $60 000 to $74 999
* 'B19001_013' - $75 000 to $99 999
* 'B19001_014' - $100 000 to $124 999
* 'B19001_015' - $125 000 to $149 999
* 'B19001_016' - $150 000 to $199 999
* 'B19001_017' - $200 000 or more

For our purposes, we are only interested in incomes at the threshold determined by Redwood City's relocation assistance policy. On the city's [renter protections webpage](https://www.redwoodcity.org/departments/community-development-department/planning-housing/renter-protections) under "Relocation Assistance Highlights" it says:

> Eligible residential households include a displaced residential household whose annual income **does not exceed eighty percent of the area median household income** for San Mateo County as adjusted for household size according to the United States Department of Housing and Urban Development and whose rental payments to the landlord remain current through the date of displacement. [^1]

The city's AMI limits can be found [here](https://www.redwoodcity.org/home/showdocument?id=16686). Income limits are also determined by household size. 

The average household size in Redwood City (according to ACS 2018 5-year data) was 2.8 persons. For the purposes of this EDA analysis, we will generalize to a household size of 3 across the entire city. This would generalize 80% AMI to $105,700.

The ACS does not have household incomes with such precision, but we will generalize 80% AMI to $100,000. Create a new variable that counts the number of households making below $100,000 and the proportion of eligibility.

```{r}
acs_rwc2 <- 
  acs_rwc %>%
  mutate(
    eligible_ami = 
      rowSums(
        select(., matches("00[2-9]$|01[1-3]$"))
      ),
    eligible_ami_prop =
      eligible_ami / B19001_001
  )
```

With this data, we can find out what proportion of Redwood City households are eligible for tenant relocation assistance upon being evicted. Keep in mind that the policy also has eligibility requirements by number of dwelling units in the building, so this proportion should be considered a relative maximum.

```{r}
weighted.mean(acs_rwc2$eligible_ami_prop, w = acs_rwc2$B19001_001)
```


Join ACS data to block group boundaries. This makes the ACS data a spatial object. Then turn it into an sf object.

```{r}
acs_rwc_sf <- 
  acs_rwc2 %>% 
  left_join(
    boundaries_cbg_city_redwood %>% select(`GEOID`, geometry),
    by = "GEOID"
  ) %>% 
  st_as_sf()
```

Let's visualize ...

```{r}
acs_rwc_sf %>% 
  ggplot() +
  geom_sf(aes(fill = eligible_ami_prop)) +
  scale_fill_gradientn(colors = RColorBrewer::brewer.pal(n = 9, name = "PuRd")) +
  geom_sf(data = bayarea_water %>% st_filter(boundaries_city_redwood, .predicate = st_intersects), fill = "aliceblue") +
  geom_sf(data = boundaries_city_redwood, color = "black", fill = NA, size = 0.2) +
  coord_sf(
    xlim = c(-122.31, -122.14),
    ylim = c(37.43, 37.57)
  )
```

For a more accurate depiction of eligiblity, we should include the other restrictions as stated on the website:

> Rental units not included in the City's ordinance include a room or any portion of a residential unit which is occupied by the landlord, a mobile home, properties of **four or fewer dwelling units** located on one lot, including single-family, duplex, tri-plex, or four-plex homes and accessory dwellings, mobile homes or housing accommodation in hotels, motels, etc. 

Create a universe of residential properties in RWC and count their land uses.
```{r}
cl_city_redwood_res <- 
  cl_city_redwood %>% 
  filter(property_indicator %in% c(10, 11)) %>% 
  left_join(data_dic_luse, by = c("land_use" = "luse_code"))

cl_city_redwood_res
```


