---
title: "CoreLogic EDA"
author: "Kate Ham"
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
# Libraries
library(tidyverse)
library(lubridate)
library(sf)

# Parameters
corelogic <- read_rds(here::here("c01-own/data-raw/all9counties.rds"))
best_vars <-   read_rds(here::here("c01-own/data/best_vars.rds"))
  # Property Indicator filter
RESIDENTIAL <- c(0:199)
  # recode function
recode_county <- function(county_fips) {
  county_fips %>% 
    recode(
      '06001' = "alameda",
      '06013' = "contra_costa", 
      '06041' = "marin",
      '06055' = "napa", 
      '06075' = "san_francisco",
      '06081' = "san_mateo",
      '06085' = "santa_clara",
      '06095' = "solano",
      '06097' = "sonoma" 
    )
}
#===============================================================================

# Code
```

## Filter for residential properties and best vars
```{r}
cl_select <-
  corelogic %>%
  select(best_vars) %>% 
  cl_select %>% filter(LAND_USE %in% RESIDENTIAL)

rm(corelogic)
```

### Missing Values in Sales Variables
```{r}
cl_select %>%
  summarize_at(vars(DOCUMENT_TYPE, RECORDING_DATE, SALE_CODE, SALE_PRICE, TRANSACTION_TYPE), ~ sum(is.na(.)))
```
GD = Grant Deed
NA = missing
DD = Deed
CR = Corporate Deed
IV = Individual Grant Deed
QC = Quit Claim Deed
WD = Warranty Deed
AF = Affidavit (of death)
IT = Interspousal Deed Transfer
JT = Joint Tenancy Deed

RECORDING_DATE missing = 234091

Remember to investigate YEAR_BUILT

```{r}
cl_select %>% 
  filter(is.na(TRANSACTION_TYPE), !is.na(DOCUMENT_TYPE))
```
Use TRANSACTION_TYPE, though there are some observations with information from DOCUMENT_TYPE, therefore with property value information.

```{r}
cl_select %>% 
  count(TRANSACTION_TYPE, sort = TRUE)
```
Most transactions were resale (1) or new construction (3).

#### Municipal/Nonprofit Properties?
```{r}
cl_select %>% 
  filter(!is.na(TRANSACTION_TYPE)) %>% 
  filter(str_detect(OWNER1_LAST_NAME, "AFFORDABLE|CITY OF|COUNTY"))
```


```{r}
cl_select <- 
  cl_select %>% 
  mutate(RECORDING_DATE = ymd(RECORDING_DATE)) %>% 
  filter(RECORDING_DATE >= ymd(20000101))
```

can also look at whole timeline to see the post-2008 spike in sales
```{r}
cl_select %>% 
  filter(RECORDING_DATE >= ymd(20100101)) %>% 
  mutate(county = recode_county(FIPS_CODE)) %>% 
  ggplot(aes(RECORDING_DATE, color = county)) +
  geom_freqpoly(binwidth = 100)
```
Looks like San Mateo, Alameda, and Contra Costa have the most sales transactions by far. Napa and Marin are the lowest.

### Which neighborhoods in San Mateo have highest sales count?
```{r}
cl_select_smc <- 
  cl_select %>% 
  mutate(county = recode_county(FIPS_CODE)) %>%
  filter(county == "san_mateo", RECORDING_DATE >= ymd(20100101))

cl_select_smc %>%
  mutate(
    lat = PARCEL_LEVEL_LATITUDE__2_6_,
    long = PARCEL_LEVEL_LONGITUDE__3_6_
  ) %>% 
  drop_na(lat, long) %>% 
  st_as_sf(coords = c("long", "lat"), crs = 4326, agr = "constant") %>% 
  ggplot() +
  geom_sf(size = 1, shape = 23, fill = "darkred")
```


```{r}
```


```{r}
```


```{r}
