---
title: "Corporate Landlords EDA"
author: "Kate Ham"
date: "7/17/2020"
output: github_document
---

# Preface

Idea for this EDA came from Tenants Together's manual guide for corporate landlord ownership lookup [pdf](https://www.tenantstogether.org/sites/tenantstogether.org/files/Wall_St_Landlord_Participatory_Action_Research_Guide.pdf). In short, the guide says to find a landlord's website and get all the listing addresses. Then, go to a proptech data source like Property Radar to check ownership information.

This EDA tries to speed up this process by scraping all the listing addresses from a landlord site with ['rvest'](https://github.com/tidyverse/rvest), then using dyplr with pre-downloaded CoreLogic data to quickly get owner information.

Limitations: The CoreLogic data downloaded is from 2019 and is only for the 9-county Bay Area. In addition, this analysis will exclusively look at [Avalon's properties](https://www.avaloncommunities.com/northern-california).

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## CSS Selector

Use the [SelectorGadget](https://selectorgadget.com/) Chrome extension to get the CSS selector for the addresses. Save the result as "css_selector".

```{r}
## Libraries
library(tidyverse)
library(rvest)

## Parameters
url_data <- "https://www.avaloncommunities.com/northern-california" # Accessed 7-17-2020
css_selector <- ".address"
```

# Webscraping

Note: If the html elements are in lists, check out this [thread](https://stackoverflow.com/questions/52650604/scrapping-li-elements-with-rvest).

```{r read}
address_data <- 
  url_data %>% 
  read_html() %>% 
  html_nodes(css = css_selector) %>% 
  html_text()
```

Next, save the address data in a tibble to make it easier to manage.

```{r, addresses}
avalon <- 
  tibble(
    owner = "avalon",
    address = address_data,
    access_date = Sys.Date()
  )
```

# CoreLogic

