---
title: "Mountain View"
author: "Kate Ham"
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
# Libraries
library(tidyverse)
library(readxl)
library(sf)
library(leaflet)
library(leaflet.extras)

# Parameters
  # ACS data
file_acs <- here::here("c01-own/data/acs/acs_mountain_view.rds")

vars_units <- 
  vars(
    B25127_048, 	
    B25127_049, 
    B25127_050, 	
    B25127_051, 	
    B25127_055, 
    B25127_056, 	
    B25127_057, 	
    B25127_058, 
    B25127_062, 
    B25127_063, 	
    B25127_064, 	
    B25127_065, 	
    B25127_069,    
    B25127_070, 	
    B25127_071, 	
    B25127_072, 	
    B25127_076, 
    B25127_077, 	
    B25127_078, 	
    B25127_079, 
    B25127_083, 	
    B25127_084, 	
    B25127_085, 
    B25127_086	
  )

vars_ami <- 
  vars(
    B25118_015,
    B25118_016,
    B25118_017,
    B25118_018,
    B25118_019,
    B25118_020,
    B25118_021,
    B25118_022,
    B25118_023
  )
#===============================================================================

# Code
```

## Section 1
```{r}
acs_mv <- 
  file_acs %>%
  read_rds() %>% 
  transmute(
    GEOID,
    total_hh = B25118_014,
    eligible_ami =
      pmap_dbl(
        select_at(st_drop_geometry(.), vars_ami),
        ~ sum(c(...))),
    eligible_ami_prop =
      eligible_ami / total_hh,
    eligible_units =
      pmap_dbl(
        select_at(st_drop_geometry(.), vars_units),
        ~ sum(c(...))),
    eligible_units_prop =
      eligible_units / total_hh,
    eligible_prop_avg =
      eligible_units_prop * eligible_ami_prop,
    eligible_prop_max =
      pmin(eligible_ami_prop, eligible_units_prop),
    eligible_prop_min =
      (eligible_ami_prop + eligible_units_prop - 1) %>% pmax(0)
  )

```

## EDA on eligibility range
```{r}
acs_mv %>% 
  mutate(GEOID = fct_reorder(GEOID, -eligible_prop_avg)) %>% 
  ggplot(aes(GEOID)) +
  geom_errorbar(aes(ymin = eligible_prop_min, ymax = eligible_prop_max)) +
  geom_point(aes(y = eligible_prop_avg), color = "red") +
  theme(
    axis.text.x = element_blank()
  )

acs_mv %>% 
  mutate(GEOID = fct_reorder(GEOID, -eligible_prop_avg)) %>% 
  ggplot(aes(x = factor(1))) +
  geom_boxplot(aes(y = eligible_prop_avg, ymax = eligible_prop_max)) 

acs_mv %>% 
  summarize(
    total_hh = sum(total_hh),
    eligible_ami = sum(eligible_ami),
    eligible_ami_prop =
      eligible_ami / total_hh,
    eligible_units = sum(eligible_units),
    eligible_units_prop =
      eligible_units / total_hh,
    eligible_prop_avg =
      eligible_units_prop * eligible_ami_prop,
    eligible_prop_max =
      pmin(eligible_ami_prop, eligible_units_prop),
    eligible_prop_min =
      (eligible_ami_prop + eligible_units_prop - 1) %>% pmax(0)
  ) %>% 
  ggplot(aes(factor(1))) +
  geom_errorbar(aes(ymin = eligible_prop_min, ymax = eligible_prop_max)) +
  geom_point(aes(y = eligible_prop_avg), color = "red") +
  theme(
    axis.text.x = element_blank()
  )
```

```{r}
pal <- 
  colorBin(
    palette = "Purples", 
    domain = acs_mv$eligible_prop_avg, 
    bins = 5
  )

labels <- 
  str_glue(
    "{round(acs_mv$eligible_prop_avg, digits = 2) * 100}", 
    "% eligible"
    )

acs_mv %>%   
  leaflet %>%
  addProviderTiles(providers$CartoDB.Positron) %>% 
  addPolygons(
    fillColor = ~ pal(eligible_prop_avg),
    color = "white",
    dashArray = 2,
    fillOpacity = 0.8,
    label = labels,
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px")
      ),
    highlight = highlightOptions(
      weight = 3,
      color = "#666",
      dashArray = "",
      fillOpacity = 0.7,
      bringToFront = TRUE
    )
  )
```

Notes:
* Avg renter household size in Mt View = 2.3 (ACS 2017 5y)
* 120% AMI, 2 person household (2018) = $120,180
  * rounded to $100k
* more than 2 units (excluding mobilehomes and condos)
  * rounded to 1 unit; condos = ?


* 'B25127_045' - Estimate Total Renter occupied	
* 'B25127_046' - Built 2010 or later	
* 'B25127_047' - Built 2010 or later 1, detached or attached	
* 'B25127_048' - Built 2010 or later 2 to 4	
* 'B25127_049' - Built 2010 or later 5 to 19
* 'B25127_050' - Built 2010 or later 20 to 49	
* 'B25127_051' - Built 2010 or later 50 or more	
* 'B25127_052' - Built 2010 or later Mobile home, boat, RV, van, etc.	
* 'B25127_053' - Built 2000 to 2009	
* 'B25127_054' - Built 2000 to 2009 1, detached or attached	
* 'B25127_055' - Built 2000 to 2009 2 to 4	
* 'B25127_056' - Built 2000 to 2009 5 to 19	
* 'B25127_057' - Built 2000 to 2009 20 to 49	
* 'B25127_058' - Built 2000 to 2009 50 or more	
* 'B25127_059' - Built 2000 to 2009 Mobile home, boat, RV, van, etc.
* 'B25127_060' - Built 1980 to 1999	
* 'B25127_061' - Built 1980 to 1999 1, detached or attached	
* 'B25127_062' - Built 1980 to 1999 2 to 4	
* 'B25127_063' - Built 1980 to 1999 5 to 19	
* 'B25127_064' - Built 1980 to 1999 20 to 49	
* 'B25127_065' - Built 1980 to 1999 50 or more	
* 'B25127_066' - Built 1980 to 1999 Mobile home, boat, RV, van, etc.	
* 'B25127_067' - Built 1960 to 1979	
* 'B25127_068' - Built 1960 to 1979 1, detached or attached	
* 'B25127_069' - Built 1960 to 1979 2 to 4
* 'B25127_070' - Built 1960 to 1979 5 to 19	
* 'B25127_071' - Built 1960 to 1979 20 to 49	
* 'B25127_072' - Built 1960 to 1979 50 or more	
* 'B25127_073' - Built 1960 to 1979 Mobile home, boat, RV, van, etc.	
* 'B25127_074' - Built 1940 to 1959	
* 'B25127_075' - Built 1940 to 1959 1, detached or attached	
* 'B25127_076' - Built 1940 to 1959 2 to 4	
* 'B25127_077' - Built 1940 to 1959 5 to 19	
* 'B25127_078' - Built 1940 to 1959 20 to 49	
* 'B25127_079' - Built 1940 to 1959 50 or more
* 'B25127_080' - Built 1940 to 1959 Mobile home, boat, RV, van, etc.	
* 'B25127_081' - Built 1939 or earlier	
* 'B25127_082' - Built 1939 or earlier 1, detached or attached	
* 'B25127_083' - Built 1939 or earlier 2 to 4	
* 'B25127_084' - Built 1939 or earlier 5 to 19	
* 'B25127_085' - Built 1939 or earlier 20 to 49	
* 'B25127_086' - Built 1939 or earlier 50 or more	
* 'B25127_087' - Built 1939 or earlier Mobile home, boat, RV, van, etc.


* 'B25118_014' - Total Renter occupied
* 'B25118_015' - Total Renter occupied Less than $5,000	
* 'B25118_016' - Total Renter occupied $5,000 to $9,999	
* 'B25118_017' - Total Renter occupied $10,000 to $14,999
* 'B25118_018' - Total Renter occupied $15,000 to $19,999
* 'B25118_019' - Total Renter occupied $20,000 to $24,999
* 'B25118_020' - Total Renter occupied $25,000 to $34,999
* 'B25118_021' - Total Renter occupied $35,000 to $49,999
* 'B25118_022' - Total Renter occupied $50,000 to $74,999
* 'B25118_023' - Total Renter occupied $75,000 to $99,999
* 'B25118_024' - Total Renter occupied $100,000 to $149,999
* 'B25118_025' - Total Renter occupied $150,000 or more
