---
title: "ACS TRA Policy Mapping"
author: "Kate Ham"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
---

This report examines the eligibility requirements of various tenant relocation assistance (TRA) policies in the San Francisco Bay Area. Tenant advocates have concerns that the TRA policies do not have significant coverage. In this report, I examine two main eligibility requirements: units and percentage of area median income (AMI).

The units requirement generally exempts smaller homes, such as single-family residences and duplexes. The AMI percentage requirement, also known as *means-testing* in essence determines which households are "poor enough" to qualify for TRA.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
# Libraries
library(tidyverse)
library(readxl)
library(sf)
library(leaflet)
library(leaflet.extras)

# Parameters
cities <- 
  c(
    "alameda",
    "san_leandro",
    "emeryville",
    "oakland",
    "berkeley",
    "san_francisco",
    "east_palo_alto",
    "redwood_city",
    "menlo_park",
    "mountain_view",
    "palo_alto",
    "san_jose"
  )
```

Load in the parameters from the `data/acs_vars` folder which has various .rds files. For details and methodology, please see `acs_vars.md`.

```{r message=FALSE, warning=FALSE}
# Parameters cont'd
filenames <- fs::dir_ls(here::here("c01-own/data/acs_vars"))
names <- str_match_all(filenames, "(\\w*_\\w*)\\.rds$")
for(i in names){
  filepath <- file.path(here::here("c01-own/data/acs_vars"), i[1])
  assign(i[2], read_rds(filepath))
}
```

The following chunk outputs maps for each individual city. I create variables for each census tract with the proportion of households eligible by both AMI and units, then provide an estimate for eligibility with `eligible_prop_avg` by multiplying the proportions.

```{r message=FALSE, warning=FALSE}

city_eligibility <- function(city) {
  data <- 
    str_glue(here::here("c01-own/data/acs/acs_"), {{city}}, ".rds") %>%
    read_rds() %>% 
    transmute(
      GEOID,
      city = {{city}},
      total_hh = B25118_014,
      eligible_ami =
        pmap_dbl(
          select_at(st_drop_geometry(.), vars_ami1),
          ~ sum(c(...))),
      eligible_ami_prop =
        eligible_ami / total_hh,
      eligible_units =
        pmap_dbl(
          select_at(st_drop_geometry(.), recode_units(city = {{city}})),
          ~ sum(c(...))),
      eligible_units_prop =
        eligible_units / total_hh,
      eligible_prop_avg =
        eligible_units_prop * eligible_ami_prop,
      eligible_prop_max =
        pmin(eligible_ami_prop, eligible_units_prop),
      eligible_prop_min =
        (eligible_ami_prop + eligible_units_prop - 1) %>% pmax(0)
    ) 
  
  assign(city, data, envir = .GlobalEnv)
  
  pal <- 
    colorBin(
      palette = "Purples", 
      domain = data$eligible_prop_avg, 
      bins = 5
    )
  
  labels <- 
    str_glue(
      "{round(data$eligible_prop_avg, digits = 2) * 100}", 
      "% eligible"
    )
  
  data %>%   
    st_transform(crs = "+proj=longlat +datum=WGS84") %>% 
    leaflet %>%
    addProviderTiles(providers$CartoDB.Positron) %>% 
    addPolygons(
      fillColor = ~ pal(eligible_prop_avg),
      color = "white",
      dashArray = 2,
      fillOpacity = 0.8,
      label = labels,
      labelOptions = labelOptions(
        style = list("font-weight" = "normal", padding = "3px 8px")
      ),
      highlight = highlightOptions(
        weight = 3,
        color = "#666",
        dashArray = "",
        fillOpacity = 0.7,
        bringToFront = TRUE
      )
    ) %>% 
    addLegend(
      pal = pal, 
      values = ~ eligible_prop_avg * 100, 
      opacity = 0.7, 
      title = "Proportion Eligible",
      labFormat = labelFormat(suffix = "%"),
      position = "bottomright"
    ) %>% 
    addControl(
      html = data$city[1] %>% str_replace("_", " ") %>% str_to_title(),
      position = "topleft"
    )
}

#===============================================================================

map(cities, city_eligibility)
```

The following chunk adds all the cities to one tibble.

```{r}
# can't use bind_rows() to preserve geometry
cities_all <- 
  rbind(
    alameda,
    san_leandro,
    emeryville,
    oakland,
    berkeley,
    san_francisco,
    east_palo_alto,
    redwood_city,
    menlo_park,
    mountain_view,
    palo_alto,
    san_jose
  ) %>% 
  st_transform(crs = "+proj=longlat +datum=WGS84")

```

The following is a plot of the error in the `eligible_prop_avg` measure since it used an average of the two proportions. Using `eligible_prop_min` and `eligible_prop_max` the following plot displays the range of error for the eligibilities.

```{r}
cities_all %>% 
  group_by(city) %>% 
  summarize(
    total_hh = sum(total_hh),
    eligible_ami = sum(eligible_ami),
    eligible_ami_prop =
      eligible_ami / total_hh,
    eligible_units = sum(eligible_units),
    eligible_units_prop =
      eligible_units / total_hh,
    eligible_prop_avg =
      eligible_units_prop * eligible_ami_prop,
    eligible_prop_max =
      pmin(eligible_ami_prop, eligible_units_prop),
    eligible_prop_min =
      (eligible_ami_prop + eligible_units_prop - 1) %>% pmax(0)
  ) %>% 
  ggplot(aes(fct_reorder(city, eligible_prop_avg))) +
  geom_errorbar(aes(ymin = eligible_prop_min, ymax = eligible_prop_max)) +
  geom_point(aes(y = eligible_prop_avg), color = "red") +
  coord_flip()

```

The following plot is a combined leaflet map of all selected cities.

```{r}
pal <- 
  colorBin(
    palette = "Reds", 
    domain = cities_all$eligible_prop_avg, 
    bins = 5
  )

labels <- 
  str_glue(
    "{round(cities_all$eligible_prop_avg, digits = 2) * 100}", 
    "% eligible"
  )

cities_all %>%   
  leaflet %>%
  addProviderTiles(providers$CartoDB.Positron) %>% 
  addPolygons(
    fillColor = ~ pal(eligible_prop_avg),
    weight = 1,
    color = "white",
    dashArray = 2,
    fillOpacity = 0.8,
    label = labels,
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px")
    ),
    highlight = highlightOptions(
      weight = 3,
      color = "#666",
      dashArray = "",
      fillOpacity = 0.7,
      bringToFront = TRUE
    )
  ) %>% 
  addLegend(
    pal = pal, 
    values = ~ eligible_prop_avg, 
    opacity = 0.7, 
    title = NULL,
    position = "bottomright"
  )
```

Note that the TRA policies have relatively and absolutely poor coverage in peninsula cities.
